Contextuality and Nonlocality in Discrete Quantum Theory

### 2015-12-16 DQC3.tex
Clean up comments in 4.2 Unitary and Hermitian operators in DQT

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Critical Analysis of the Measurement Process in CQT}

@@

After we define the state vectors and the Hermitian dot product, we
are ready to define unitary and Hermitian operators representing time
evolutions and observables, respectively. All unitary operators can
represent time-evolution operators. Because a unitary operator~$\mathbf{U}$
preserves norm, a physical state~$\ket{\Psi}$ always evolves to
another physical state~$\mathbf{U}\ket{\Psi}$ according to $\mathbf{U}$.\footnote{Although some unitary operators might not be induced by a Hamiltonian,
we might discuss further in our upcoming papers.}

@@

3, the measurement result of an observable~$\mathbf{O}$ is a real
number and a eigenvalue of $\mathbf{O}$; the post-measurment state
is a physical state as well as an eigenvector of $\mathbf{O}$. However,
a Hermitian operator~$\mathbf{H}$ over $\ff{p^{2}}$ may not be
qualified for an observable because $\mathbf{H}$ may not have suitable
eigenvalues and eigenvectors to define the measurement precess, or
equavalently $\mathbf{H}$ is not unitarily diagonalizable, for the
following reasons:

@@

\item \label{enu:eigenvalueCount}$\mathbf{H}$ may lack of eigenvalues.
\\
\yutsung{No eigenvalue is a problem, but how to clearly say why no
enough eigenvalue is a problem?}\\
Formally, the total multiplicity of eigenvalues of $\mathbf{H}$ in
$\ff{p^{2}}$ may be less than $d$, where the multiplicity of an
eigenvalue $\lambda$ is $K_{\lambda}$ means that $\left(x-\lambda\right)^{K_{\lambda}}$

@@

\det\left(\mathbf{A}-x\mathds{1}\right)\textrm{ ,}\label{eq:characteristic}

@@

the characteristic polynomial. This happens because $\ff{p^{2}}$
is not algebraically closed~\cite{DummitFoote2004} so that the solutions
of the characteristic polynomial may be in an proper extension field
of $\ff{p^{2}}$. For example, consider a Hermitian matrix 

@@

which have no solution in $\ff{9}$ so that their total multiplicities
are 0 less than 3. In this case, $\mathbf{H}_{1}$ should not be an
observable because there is no eigenvalue as any measurement result. 
\item \label{enu:degeneracy}\yutsung{The same. How to say clearly why
no enough eigenvector is a problem?}\\
 The dimension of an eigenspace may be less than the multiplicity
of its eigenvalue. For example, consider a Hermitian matrix 

@@

1 & -1+\rmi\end{array}\right)^{T}$. The dimension of eigenspace is 1 less than 2 so that $\mathbf{H}_{2}$

@@

observable, $\mathbf{H}_{2}$ would always have the outcome of measurement

@@

1 & -1+\rmi\end{array}\right)^{T}$. However, only operator always producing $1$ is the identity operator
which change no state. Therefore, $\mathbf{H}_{2}$ may not be intuitively
as an observable.\\
 In order to understand this non-diagonalability, we can diagonalize
them in $\C$. The eigenvalues of $\mathbf{H}_{2}$ in $\C$ are $1$
and $-2$. 

@@

-2 & \equiv & 1\left(\bmod p\right)

@@

with dimension $1$ in $\C$ map to one eigenspace still with dimension
$1$ in $\ff{9}$. 

@@

are $-1-\rmi$ with $\left(\begin{array}{cc}
1 & -1-\rmi\end{array}\right)^{T}$ and $-1+\rmi$ with $\left(\begin{array}{cc}
1 & 1+\rmi\end{array}\right)^{T}$. It is easy to verify the eigenvalues are not in $\ff{3}$, and the

@@

 If we diagonalize $\mathbf{H}_{3}$ in $\C$, we have the eigenvalues
of $\mathbf{H}_{3}$ in $\C$ are $-1\pm\sqrt{2}$. If we extend the
modulo $p$ to algebraic number by 

@@

a\equiv b\left(\bmod p\right)\Leftrightarrow\pm\sqrt{a}\equiv\pm\sqrt{b}\left(\bmod p\right)

@@

-1\pm\sqrt{2} & \equiv & -1\pm\rmi\left(\bmod p\right)

@@

In another word, although the extended modulo $p$ operation respects
any field operations and roots, it does not respect the complex conjugate
so that any definition depend on the complex conjugate may not be
preserved when we map them into finite fields. \yutsung{Although
may not need to type it down, need to check the definition of the
map, passing congruence into square root, and partial homomorphism...
etc.}

@@

\yutsung{
I don't know whether there is a relation between the occurrence of
zero norm post-measurement states and anomalies in the list of possible
observables, especially we have already required any observables must
be unitarily diagonalizable.

I am sure I have asked Prof. Lunts ``what we can say for the eigenvalues
of Hermitian matrices over finite fields,'' and he seems not sure.
I guess I also asked Prof. Thurston the same question, but I am not
sure...
}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{\label{sec:Critical-Analysis-CQT}Critical Analysis of the Measurement
Process in CQT}

@@

After we define the state vectors and the Hermitian dot product, we
are ready to define unitary and Hermitian operators representing time
evolution and observables, respectively. All unitary operators can
represent time-evolution operators. Because a unitary operator~$\mathbf{U}$
preserves norm, a physical state~$\ket{\Psi}$ always evolves to
another physical state~$\mathbf{U}\ket{\Psi}$ according to $\mathbf{U}$.\footnote{Although some unitary operators might not be induced by a Hamiltonian,
we might discuss further in our upcoming papers.}

@@

\ref{sec:Critical-Analysis-CQT}, the measurement result of an observable~$\mathbf{O}$
is a real eigenvalue of $\mathbf{O}$; the post-measurment state is
a physical eigenstate of $\mathbf{O}$. However, a Hermitian operator~$\mathbf{H}$
over $\ff{p^{2}}$ may not be qualified for an observable because
$\mathbf{H}$ may not have suitable eigenvalues and eigenvectors to
define the measurement process, or equivalently $\mathbf{H}$ is not
unitarily diagonalizable, for the following reasons: 

@@

\item \label{enu:eigenvalueCount}$\mathbf{H}$ may lack eigenvalues. More
precisely, the total multiplicity of eigenvalues of $\mathbf{H}$
in $\ff{p^{2}}$ may be less than $d$, where $K_{\lambda}$ is the
multiplicity of an eigenvalue $\lambda$ means that $\left(x-\lambda\right)^{K_{\lambda}}$

@@

\det\left(\mathbf{H}-x\mathds{1}\right)\textrm{ ,}\label{eq:characteristic}

@@

the characteristic polynomial. In another word, $\ff{p^{2}}$ is not
algebraically closed~\cite{DummitFoote2004}. For example, consider
a Hermitian matrix 

@@

which have no solution in $\ff{9}$ so that the total multiplicities
is 0 less than 3. Hence, $\mathbf{H}_{1}$ should not be an observable
because there is no eigenvalue as any measurement result. 
\item \label{enu:degeneracy}The dimension of an eigenspace may be less
than the multiplicity of its eigenvalue. For example, consider a Hermitian
matrix 

@@

1 & -1+\rmi\end{array}\right)^{T}$. The dimension of the eigenspace is 1 less than 2 so that $\mathbf{H}_{2}$

@@

observable, $\mathbf{H}_{2}$ would always have the measurement result

@@

1 & -1+\rmi\end{array}\right)^{T}$. In another word, if we measured $\mathbf{H}_{2}$ by the Stern-Gerlach
experiment, we would always produce a $\left(\begin{array}{cc}
1 & -1+\rmi\end{array}\right)^{T}$ beam without even splitting the input beam. This violates our intuitively
of quantum measurement. Therefore, $\mathbf{H}_{2}$ should not be
an observable.\\
In order to understand the reason why $\mathbf{H}_{2}$ is not diagonalizable,
we can diagonalize $\mathbf{H}_{2}$ in $\C$. The eigenvalues of
$\mathbf{H}_{2}$ in $\C$ are $1$ and $-2$. 

@@

-2 & \equiv & 1\left(\bmod3\right)

@@

with dimension $1$ in $\C$ map to the same eigenspace still with
dimension $1$ in $\ff{9}$. 

@@

are $-1+\rmi$ with $\left(\begin{array}{cc}
1 & 1+\rmi\end{array}\right)^{T}$ and $-1-\rmi$ with $\left(\begin{array}{cc}
1 & -1-\rmi\end{array}\right)^{T}$. It is easy to verify the eigenvalues are not in $\ff{3}$, and the

@@

If we diagonalize $\mathbf{H}_{3}$ in $\C$, $\mathbf{H}_{3}$ has
the eigenvalues $-1\pm\sqrt{2}$. If we extend the modulo $p$ to
some algebraic numbers by 

@@

a\equiv b\left(\bmod p\right)\Leftrightarrow\pm\sqrt{a}\equiv\pm\sqrt{b}\left(\bmod p\right)\textrm{ ,}

@@

-1\pm\sqrt{2} & \equiv & -1\pm\rmi\left(\bmod3\right)\textrm{ ,}

@@

This means the real eigenvalues in $\C$ may be corresponding to the
complex eigenvalues in $\ff{p^{2}}$. The complex eigenvalues in $\ff{p^{2}}$
is further related to the zero-norm eigenstates because if an eigenstate
is not zero-normed, its corresponding eigenvalue must be in $\ff{p}$.\\
\yutsung{Although may not need to type it down, need to check the
definition of the map, passing congruence into square root, and partial
homomorphism... etc.}

@@
